---
title: "ExPra_fetch_participantData"
author: "Noah Rischert"
date: "2025-04-01"
output:
  word_document: default
  html_document:
    code_folding: hide
  pdf_document: default
---
# prinz hello worl
```{r setup, include=FALSE}
# Set global chunk options
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE)
```

## Introduction

This R Markdown document demonstrates how to:

1. Install and load the **smartr** package from GitHub,  
2. Set your working directory, and  
3. Use the `get_JATOS_data()` function to fetch experiment data from a JATOS server.


!! NOTE: You will need to be either at UZH or on the UZH VPN to access the data server !!

---

## 1. Install (or Update) the Required Packages

Here, we install **smartr** directly from GitHub using `devtools::install_github()`.

```{r install-packages}
# If 'devtools' is not installed, uncomment and run the line below:
# install.packages("devtools")

# Install the 'smartr' package from GitHub. This package will allow us to fetch the data from the JATOS server.
# When installing the smartr package, please select "3" (None) for the updates and "no" for installing from sources the packages which need compilation.

# The package smartr was updated. We can therefore:
#remove.packages("smartr")  # remove any old installation

#restart R completely ("Session>Restart R" or just close and reopen rstudio)

#Install the packages again. Once is enough, you don't need to do this every time you run the this file.
#devtools::install_github("chenyu-psy/smartr@develop")


# Tydiverse is a very useful package for data analysis. We will look at it next week.
#install.packages("tidyverse")

# Load the package into your session
library(smartr)
library(tidyverse)

```

---

## 2. Set Your Working Directory

To make our workflow more efficient, we need to make sure that our working directory makes sense.
Below is one approach to ensure your working directory is set to the folder containing this R Markdown file.  
When we're using RStudio, `rstudioapi::getActiveDocumentContext()` can help, but note that it only works interactively, *not* when knitting.

```{r set-wd, eval=FALSE}
# Example: Set working directory to the current file's location in RStudio
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# Or manually set your working directory:
# setwd("C:/Path/To/Your/Folder")

# Check whether this worked:
getwd()
```

---

## 3. Fetch JATOS Data with `get_JATOS_data()`


In practice, replace the placeholders (`your_token`, `your_study_id`, `your_batch_id`) with your actual token and IDs.


!! NOTE: If you call to the server to grab the data while a participant is currently running the study,
the partial data will be downloaded. If you run this again in the future, R will see that you already
have downloaded this participant's data and will NOT download the new complete dataset for this participant.
Before running this code, locate the local file containing the metadata (by default, this will be a folder 
in the current directory of R with the name "ExPraData"). Inside you will see a metadata file and a JATOS folder.
Delete both of these items. !!

```{r fetch-data}
# Define your parameters
your_token <- "jap_fvOoehoOVsH0g3plDGXIuGmY743rf5d81b0ef"

your_batch_id <- "1725" # Again, this can be different for each group.

# Optionally specify a path for saving data (default is the working directory)
data_output_path <- "./ExPraData/"
if(!dir.exists(data_output_path)) dir.create(data_output_path)

# Fetch the data
jatos_data <- get_JATOS_data(
  token     = your_token,
  batchId   = your_batch_id,
  dataPath  = data_output_path,
  attachments = FALSE
)

# This function gives us 2 things: 
# A metadata file called jatos_data, which contains information about the data we have downloaded
# The actual downloaded data, at our specified location

# Print a preview of the data
head(jatos_data)

# Also make sure to look at the metadata file (jatos_data) itself.
```

### Notes on the `get_JATOS_data()` Function

- **`token`**: Create a JATOS personal access token under your user account on the JATOS server.  
- **`batchId`**: Each study can have multiple batch sessions (e.g., one for a student pool, one for external participants). Provide one or more batch IDs in a vector.  
- **`dataPath`**: Where downloaded data and metadata files will be saved (default: working directory).  

---

## 4. Using the metadata, we can filter for the data of participants that actually completed the experiment.

```{r participant-numbers}
# Filter the metadata for files that actually contain a file.
filtered_metadata <- jatos_data %>% 
  filter(!is.na(file))

# We could also filter by other criteria, for example the file size. If the file size is very very small, we know it can't contain a full dataset.
# For example:
# THIS STEP LARGLY DEPENDS ON YOUR OWN EXPERIMENT. DO NOT BLINDLY RUN IT WITHOUT CHECKING OUT YOUR FILE SIZES FIRST, OR YOU RISK REMOVING VALID EXPERIMENTAL DATA
filtered_metadata <- filtered_metadata %>%
  filter(fileSize >100)

```

---

## 5. Lets also look at a dataframe. Does it look like what we might expect?

```{r df-inspection}
# Lets just look at a single file of those we downloaded. We use the metadata to access our downloaded data, as the metadata contains all the existing file paths.
single_participant_df <- read_json_data(filtered_metadata$file[1])

# Now we can look at single_participant_df and check whether the data looks as expected.
```


## 6. Finally, we can combine the Data from All Participants

```{r df-combined}
# Now we parse *all* the 'data.txt' files for the current batch. Again, all paths are listed in the metadata file under $file.
all_participants_df <- read_json_data(filtered_metadata$file)
```

## 7. Now we can filter for the subject hour information

```{r subject hours}
survey_responses <- all_participants_df %>%
  # Keep only the survey trials
  filter(trial_type == "survey") %>%
  # We only need subject and response columns right now
  select(participant, response) %>%
  # Unnest the list column so that each response item becomes its own row
  unnest_longer(response, indices_to = "question", values_to = "answer") %>%
  # Now pivot from long to wide, making question names into columns
  pivot_wider(
    names_from  = question,
    values_from = answer
  )
```

## 8. Calculate the number of participants collected

```{r subject count}
sprintf("%i Participants Collected", length(unique(na.omit(all_participants_df$participant))))
```